---
title: "Complex matrix visualization"
description: |
  Complex matrix visualization with BraKetVue - an interactive article in Distill RMarkdown
author:
  - name: Piotr Migdał 
    url: https://p.migdal.pl
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
  - name: Klem Jankiewicz
    url: http://jankiewiczstudio.com/
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='asis', echo=FALSE, message=FALSE}
cat(
  "<script src='https://cdn.jsdelivr.net/npm/vue@2.6.12'></script>",
  "<script src='https://unpkg.com/quantum-tensors@0.4.10/dist/quantum-tensors.min.js'></script>",
  "<script src='https://unpkg.com/bra-ket-vue@0.3.1/dist/bra-ket-vue.min.js'></script>"
)
```


Some intro notes on matrix viz and its importance.
Etc, etc.

E.g. "quantum lorem ipsum" (using <https://transformer.huggingface.co/doc/gpt2-large>):

**Quantum mechanics is** a powerful, revolutionary system of ideas that are fundamentally changing our view of the universe and the universe itself. It is, to use a common term, the "singularity of science " and it has a profound impact on all aspects of our lives and on everything we do." (4) The problem with the Sing ularity is that it creates uncertainty , but uncertainty is unavoidable because the system we have set up to be able to predict what will happen in the future and the system we have set up to measure the past, does not do this well.

## BraKetVue library

<div id='matrix1'></div>

Here more notes.
And even more.

<div id='matrix2'></div>

## Equations in general

Not only we can have Vue components, but also regular equations in $\LaTeX$, e.g.:

$$\left| \psi \right\rangle = \frac{\left| 0 \right\rangle - \left| 1 \right\rangle}{\sqrt{2}}$$

## Notes

This note is created with Learn more about using [Distill for R Markdown](https://rstudio.github.io/distill).

As of now, source code is at [github.com/stared/bra-ket-vue-art](https://github.com/stared/bra-ket-vue-art).

We should ref or consider:

- [Clustergrammer in Jupyter Notebook](https://clustergrammer.readthedocs.io/clustergrammer_widget.html) (see potentially [ipywidgets.readthedocs.io](https://ipywidgets.readthedocs.io/en/stable/examples/Widget%20Custom.html))
- [htmlwidgets for R](http://www.htmlwidgets.org/)
- (found in <https://bost.ocks.org/mike/algorithms/>)

> This is not a convential rainbow color scale, which is nominally considered harmful, but a cubehelix rainbow with improved perceptual properties.

- [How The Rainbow Color Map Misleads - Robert Kosara](https://eagereyes.org/basics/rainbow-color-map)
- [A Less-Angry Rainbow - Mike Bostock](https://bl.ocks.org/mbostock/310c99e53880faec2434)
- [chapter 3. Visualization of quantum sequences of Piotr Migdał PhD Thesis](https://arxiv.org/abs/1412.6796)
- [QuTiP example: Qubism visualizations - Piotr Migdał](https://nbviewer.jupyter.org/github/qutip/qutip-notebooks/blob/master/examples/qubism-and-schmidt-plots.ipynb)
- [Visualizing complex-valued functions with Matplotlib and Mayavi - Domain coloring method](https://nbviewer.jupyter.org/github/empet/Math/blob/master/DomainColoring.ipynb)


```{js, results='asis', echo=FALSE, message=FALSE}
const { Elements, Operator, Dimension, Cx } = QuantumTensors;
const { ComplexLegend, MatrixViewer } = BraKetVue;

const operator = Elements.beamSplitter(45);
const operator2 = Operator.fromSparseCoordNames([
  ['u', 'u', Cx(1)],
  ['d', 'd', Cx(-1)],
], [Dimension.spin()]);

new Vue({
  el: "#matrix1",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
  	ComplexLegend,
    MatrixViewer
  },
  data() {
  	return {
    	operator,
      operator2
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#matrix2",
  template: "<matrix-viewer :operator-raw='operator2' :dark-mode='false' />",
  components: {
  	ComplexLegend,
    MatrixViewer
  },
  data() {
  	return {
    	operator,
      operator2
		}
  }
})
```


