---
title: "Ket vis"
description: |
  Pure states visualized with BraKetVue
author:
  - name: Piotr Migdał 
    url: https://p.migdal.pl
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
  - name: Klem Jankiewicz
    url: http://jankiewiczstudio.com/
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='asis', echo=FALSE, message=FALSE}
cat(
  "<script src='https://cdn.jsdelivr.net/npm/vue@2.6.12'></script>",
  "<script src='https://unpkg.com/quantum-tensors@0.4.10/dist/quantum-tensors.min.js'></script>",
  "<script src='https://unpkg.com/bra-ket-vue@0.3.1/dist/bra-ket-vue.min.js'></script>"
)
```

## Ket states for qubits

We start with a 3-qubit state:

<div id='ket0'></div>

Now, let's apply Hadamard gate to the first qubit.
Here is the resulting sta

<div id='ket1'></div>

Now, let's apply CNOT gate to the first two qubits:

<div id='ket2'></div>

Finally, we apply Toffoli gate to all three qubits:

<div id='ket3'></div>

## Gates

If you've ever wondered how does the Hadamard matrix look like, here's one:

<div id='gate-h'></div>

If we have three qubits, the operator acting on them is $H \otimes I \otimes I$, that is:

<div id='gate-hii'></div>

```{js, results='asis', echo=FALSE, message=FALSE}
const { Circuit, Gates } = QuantumTensors;
const { KetViewer, MatrixViewer } = BraKetVue;

const circuitHistory = [];
Circuit.qubits(3)
  .saveTo(circuitHistory)
  .H(0)
  .saveTo(circuitHistory)
  .CNOT(0, 1)
  .saveTo(circuitHistory)
  .TOFFOLI(0, 1, 2)
  .saveTo(circuitHistory);

function showState(elName, vector) {
  new Vue({
    el: elName,
    template: "<ket-viewer :vector='vector' :dark-mode='false'/>",
    components: {
      KetViewer,
    },
    data() {
  	  return { vector }
    }
  })
}

function showGate(elName, operator) {
  return new Vue({
    el: elName,
    template: "<matrix-viewer :operator-raw='operator' :dark-mode='false'/>",
    components: {
      MatrixViewer,
    },
    data() {
  	  return { operator }
    }
  })
}

showState('#ket0', circuitHistory[0].vector)
showState('#ket1', circuitHistory[1].vector)
showState('#ket2', circuitHistory[2].vector)
showState('#ket3', circuitHistory[3].vector)

console.log(Gates.H().outer(Gates.I()).outer(Gates.I()))

showGate('#gate-h', Gates.H())
showGate('#gate-hii', Gates.H().outer(Gates.I()).outer(Gates.I()))
```

```{css, echo=FALSE}
.matrix-viewer {
  margin: auto;
}
```