---
title: "Logic operations, truth tables and matrices"
description: |
  Classical logics visualized with BraKetVue
author:
  - name: Piotr Migdał 
    url: https://p.migdal.pl
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
  - name: Klem Jankiewicz
    url: http://jankiewiczstudio.com/
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='asis', echo=FALSE, message=FALSE}
cat(
  "<script src='https://cdn.jsdelivr.net/npm/vue'></script>",
  "<script src='https://unpkg.com/quantum-tensors'></script>",
  "<script src='https://unpkg.com/bra-ket-vue'></script>"
)
```


Logic operations are basic building block of any more complicated computing.


## Simple operations

### NOT

$$ \neg a $$

<div id='op-not'></div>




### AND

$$ a \wedge b $$

<div id='op-and'></div>

### OR

$$ a \vee b $$

### XOR

### NAND



## Equations in general

Not only we can have Vue components, but also regular equations in $\LaTeX$, e.g.:

$$\left| \psi \right\rangle = \frac{\left| 0 \right\rangle - \left| 1 \right\rangle}{\sqrt{2}}$$

## Reversibility

## Notes

This note is created with Learn more about using [Distill for R Markdown](https://rstudio.github.io/distill).

As of now, source code is at [github.com/stared/bra-ket-vue-art](https://github.com/stared/bra-ket-vue-art).

This particualr post inspired by [Kronecker Product In Circuits by Lei Mao ](https://leimao.github.io/blog/Kronecker-Product-In-Circuits/) and [Quantum Entanglement and Superdense Coding by Victory Omole](https://vtomole.com/blog/2018/03/03/sd).


```{js, results='asis', echo=FALSE, message=FALSE}
const { Elements, Operator, Dimension, Cx } = QuantumTensors;
const { ComplexLegend, MatrixViewer } = BraKetVue;

const opNot = Operator.fromSparseCoordNames([
  ['1', '0', Cx(1)],
  ['0', '1', Cx(1)],
], [Dimension.position(2, 'bit')]);

new Vue({
  el: "#op-not",
  template: "<matrix-viewer :operator-raw='opNot' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
    	opNot
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
const opAnd = Elements.beamSplitter(45);

new Vue({
  el: "#op-and",
  template: "<matrix-viewer :operator-raw='opAnd' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      opAnd
		}
  }
})
```


