---
title: "Logic operations, truth tables and matrices"
description: |
  Classical logics visualized with BraKetVue
author:
  - name: Piotr Migdał 
    url: https://p.migdal.pl
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
  - name: Klem Jankiewicz
    url: http://jankiewiczstudio.com/
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='asis', echo=FALSE, message=FALSE}
cat(
  "<script src='https://cdn.jsdelivr.net/npm/vue'></script>",
  "<script src='https://unpkg.com/quantum-tensors'></script>",
  "<script src='https://unpkg.com/bra-ket-vue'></script>"
)
```


> Truth tables, reversibility and the gate (pun totally intended) to quantum computing.

> I've seen things you wouldn't believe. CNOT glitter near the Toffoli gate.

Logic operations are basic building block of any more complicated computing.
They are being used as base CPU instructions, and in high-level computing.


## Simple operations

We start with  0 and 1.

### NOT

The simplest operation is negation.

$$ \neg a $$
It takes one bit, and swaps it.

$$0 \mapsto 1 \hspace{1cm} 1 \mapsto 0$$
Or we can write the same thing using a matrix:

<div id='op-not'></div>

Note that in this matrix all values are either 1 or 0. 


### AND

$$ a \wedge b $$

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 0 \hspace{1cm} 10 \mapsto 0 \hspace{1cm} 11 \mapsto 1$$

<div id='op-and'></div>

### OR

$$ a \vee b $$

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 1$$

<div id='op-or'></div>

### XOR

$$a \oplus b $$


$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 0$$

<div id='op-xor'></div>


### NAND

$$00 \mapsto 1 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 0 \mapsto 0$$

<div id='op-nand'></div>

### COPY

$$0 \mapsto 00 \hspace{1cm} 1 \mapsto 11$$

<div id='op-copy'></div>

### FIRST

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 0 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 1$$

<div id='op-first'></div>

### SECOND

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 0 \hspace{1cm} 11 \mapsto 1$$

<div id='op-second'></div>

### SWAP

$$00 \mapsto 00 \hspace{1cm} 01 \mapsto 10 \hspace{1cm} 10 \mapsto 01 \hspace{1cm} 11 \mapsto 11$$

<div id='op-swap'></div>

### ERASE

$$0 \mapsto 0 \hspace{1cm} 1 \mapsto 0$$

<div id='op-erase'></div>


## Reversibility

Both `NOT` are `SWAP` are reversible - just apply them twice, and you get your original input.

In general, these operations are not reversible. If we know one bit there is no way to guess two bits. For the logic gates that maintain the number of bits, bioth `NOT` are `SWAP` are reversible - just apply them twice, and you get your original input. It is not the general case though - even if we have the same number of bits, we may have lost all information - as it happends with `ERASE`.

So, there is a different question - can we make a gate reversible by adding extra bits to its output? Let's check out! For operations `FIRST` and `SECOND` it is simple - we just add the missing bit. How about `XOR`?

<div id='op-xor-2'></div>

Mouseover the columns - you see that for the first output bit we get the same value as in the original `XOR`. For the second output bit, there are possible options. 

To make logic operation reversible, we need to have exactly one dot in each column and each row. That is, we not only guarantee that for each input we get a single, deterministic output, but also that for each output we can get a single, deterministic input.

Can you eliminate some values, so that we get the right combination? Think of it as playing some variant of Sudoku.

**(HERE WE NEED SOME PAUSE OR 'click to reveal')**

In fact, there are 4 different solutions. On the them is a `CNOT` gate (controlled-NOT) with  

<div id='op-cnot'></div>

So, can we reverse `AND` by adding ancilla bits? 

<div id='op-and-2'></div>

Well, the task is not harder... it is impossible. Let's see why! For the output `1` we don't need any extra information, we know that the input is `11`. Sounds good? Not so! For `0` there are three possible possibilities `00`, `01`, `11`, and we cannot encode them in a single bit.

We can give up.... or add one more ancilla bit.
So, it's time for another Sudoku:

**(This example broken, as we need to change the first bit to maintain balance. It requires other adjustments. Maybe actually this need to stay, but make add more step. This actually looks didactic.)**

<div id='op-and-3'></div>

There are much more possiblities.

**(I need to think about the best / most intuitive solution.)**

<div id='op-and-3-sol'></div>


## To do

- more explanations for simple gates
- interactive widget to multiply operations
- fix original 

## Notes

This note is created with Learn more about using [Distill for R Markdown](https://rstudio.github.io/distill).

As of now, source code is at [github.com/stared/bra-ket-vue-art](https://github.com/stared/bra-ket-vue-art).

This particualr post inspired by [Kronecker Product In Circuits by Lei Mao ](https://leimao.github.io/blog/Kronecker-Product-In-Circuits/) and [Quantum Entanglement and Superdense Coding by Victory Omole](https://vtomole.com/blog/2018/03/03/sd).


```{js, results='asis', echo=FALSE, message=FALSE}
const { Elements, Operator, Dimension, Cx } = QuantumTensors;
const { ComplexLegend, MatrixViewer } = BraKetVue;

const opNot = Operator.fromSparseCoordNames([
  ['1', '0', Cx(1)],
  ['0', '1', Cx(1)],
], [Dimension.position(2, 'bit')]);

new Vue({
  el: "#op-not",
  template: "<matrix-viewer :operator-raw='opNot' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
    	opNot
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-and",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['0', '01', Cx(1)],
          ['0', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-or",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-xor",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['0', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-nand",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['1', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['0', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-copy",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '0', Cx(1)],
          ['11', '1', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-first",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['0', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-second",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['0', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-swap",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(1)],
          ['10', '01', Cx(1)],
          ['01', '10', Cx(1)],
          ['11', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-erase",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '0', Cx(1)],
          ['0', '1', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-xor-2",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['10', '01', Cx(0.5)],
          ['10', '10', Cx(0.5)],
          ['00', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['11', '01', Cx(0.5)],
          ['11', '10', Cx(0.5)],
          ['01', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-cnot",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(1)],
          ['11', '01', Cx(1)],
          ['10', '10', Cx(1)],
          ['01', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}

new Vue({
  el: "#op-and-2",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
  	  // use outer product instead?
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['00', '01', Cx(0.5)],
          ['00', '10', Cx(0.5)],
          ['10', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['01', '01', Cx(0.5)],
          ['01', '10', Cx(0.5)],
          ['11', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
const oneBitMix = Operator.fromSparseCoordNames([
    ['0', '0', Cx(0.5)],
    ['0', '1', Cx(0.5)],
    ['1', '0', Cx(0.5)],
    ['1', '1', Cx(0.5)],
  ],
  [Dimension.position(2, 'bit')],
  [Dimension.position(2, 'bit')])

new Vue({
  el: "#op-and-3",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['00', '01', Cx(0.5)],
          ['00', '10', Cx(0.5)],
          ['10', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['01', '01', Cx(0.5)],
          ['01', '10', Cx(0.5)],
          ['11', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')]
      ).outer(oneBitMix)
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}

new Vue({
  el: "#op-and-3-sol",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['000', '000', Cx(1)],
          ['010', '010', Cx(1)],
          ['001', '100', Cx(1)],
          ['111', '110', Cx(1)],
          ['100', '001', Cx(1)],
          ['110', '011', Cx(1)],
          ['101', '101', Cx(1)],
          ['011', '111', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```