---
title: "Logic operations, truth tables and matrices"
description: |
  Classical logics visualized with BraKetVue
author:
  - name: Piotr Migdał 
    url: https://p.migdal.pl
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
  - name: Klem Jankiewicz
    url: http://jankiewiczstudio.com/
    affiliation: Quantum Flytrap
    affiliation_url: https://quantumgame.io
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, results='asis', echo=FALSE, message=FALSE}
cat(
  "<script src='https://cdn.jsdelivr.net/npm/vue'></script>",
  "<script src='https://unpkg.com/quantum-tensors'></script>",
  "<script src='https://unpkg.com/bra-ket-vue'></script>"
)
```


Truth tables, reversibility and a gate (pun totally intended) to quantum computing.

> I've seen things you wouldn't believe. CNOT glitter near the Toffoli gate.

Logic operations are basic building block of any more complicated computing.
They are being used as base CPU instructions, and in high-level programming language.

We will use matrix notation for these logical operations - partly to show [truth tables](https://en.wikipedia.org/wiki/Truth_table) in a nice and interactive way, and partly to relate it to physics (reversible computation and quantum computing).


## Simple operations

We start with 0 (for false) and 1 (for true).


### NOT

The simplest operation is negation. Depending on a language, we write it differently: 

* $\neg a$ in mathematics
* `not a` in Python
* `!a` in JavaScript

If we want to explicitly define, we need to define its value for each input:

$$0 \mapsto 1 \hspace{1cm} 1 \mapsto 0$$
It reads "0 maps to 1, 1 maps to 0". 
We can write the same thing using a matrix:

<div id='op-not'></div>

In this matrix all values are either 1 (red circles) or 0 (well, nothing).
When you mouseover an input column, you will see the output.


### AND

Conjunction ($a \wedge b$  in maths, `a and b` in Python, `a && b` in JavaScript) is true if and only if both conditions are true:


$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 0 \hspace{1cm} 10 \mapsto 0 \hspace{1cm} 11 \mapsto 1$$

<div id='op-and'></div>

### OR

Alternative ($a \vee b$ in maths, `a or b` in Python, `a || b` in JavaScript) is true if at least one condition is true.


$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 1$$

<div id='op-or'></div>

### XOR

Exclusive OR ($a \oplus b$ in maths, `a ^ b` in Python and JavaScript) means *either a or b*, and is true if exactly one condition is true.


$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 0$$

<div id='op-xor'></div>


### NAND

NAND stands for non-AND, $\neg (a \wedge b)$.

$$00 \mapsto 1 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 0 \mapsto 0$$

<div id='op-nand'></div>

We rearely use it directly (in mathematics in programming we typically combine AND and NOT). At the same time, it is [an universal logic gate](https://en.wikipedia.org/wiki/Functional_completeness), meaning that ALL other logic gates can be constructed out of it. It has a big meaning in the design of circuits. 


### COPY

There is one catch, though: we need to be able to use one output in a few inputs.
If we want to write everything as matrices, we need to be explicit about the copy operation:

$$0 \mapsto 00 \hspace{1cm} 1 \mapsto 11$$

<div id='op-copy'></div>

### FIRST

If we want to be explicit about other operations, we can take the first or the second value, discarding the other.

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 0 \hspace{1cm} 10 \mapsto 1 \hspace{1cm} 11 \mapsto 1$$

<div id='op-first'></div>

### SECOND

$$00 \mapsto 0 \hspace{1cm} 01 \mapsto 1 \hspace{1cm} 10 \mapsto 0 \hspace{1cm} 11 \mapsto 1$$

<div id='op-second'></div>

### SWAP

We may want to change the order of variables:

$$00 \mapsto 00 \hspace{1cm} 01 \mapsto 10 \hspace{1cm} 10 \mapsto 01 \hspace{1cm} 11 \mapsto 11$$

<div id='op-swap'></div>

### ERASE

Or, we may want to forget:

$$0 \mapsto 0 \hspace{1cm} 1 \mapsto 0$$

<div id='op-erase'></div>


## Reversibility

Both `NOT` are `SWAP` are reversible - just apply them twice, and you get your original input.

In general, these operations are not reversible. If we know one bit there is no way to guess two bits. For the logic gates that maintain the number of bits, both `NOT` are `SWAP` are reversible - just apply them twice, and you get your original input. It is not the general case though - even if we have the same number of bits, we may have lost all information - as it happends with `ERASE`.

So, there is a different question - can we make a gate reversible by adding extra bits to its output? Let's check out! For operations `FIRST` and `SECOND` it is simple - we just add the missing bit. How about `XOR`?

<div id='op-xor-2'></div>

Mouseover the columns - you see that for the first output bit we get the same value as in the original `XOR`. For the second output bit, there are possible options. 

To make logic operation reversible, we need to have exactly one dot in each column and each row. That is, we not only guarantee that for each input we get a single, deterministic output, but also that for each output we can get a single, deterministic input.

Can you eliminate some values, so that we get the right combination? Think of it as playing some variant of Sudoku.

**(HERE WE NEED SOME PAUSE OR 'click to reveal')**

In fact, there are 4 different solutions. On the them is a `CNOT` gate (controlled-NOT) with  

<div id='op-cnot'></div>

So, can we reverse `AND` by adding ancilla bits? 

<div id='op-and-2'></div>

Well, the task is not harder... it is impossible. Let's see why! For the output `1` we don't need any extra information, we know that the input is `11`. Sounds good? Not so! For `0` there are three possible possibilities `00`, `01`, `11`, and we cannot encode them in a single bit.

We can give up.... or add one more ancilla bit.
So, it's time for another Sudoku:

**(This example broken, as we need to change the first bit to maintain balance. It requires other adjustments. Maybe actually this need to stay, but make add more step. This actually looks didactic.)**

<div id='op-and-3'></div>

There are much more possiblities.

**(I need to think about the best / most intuitive solution.)**

<div id='op-and-3-sol'></div>


## To do

- interactive widget to multiply operations
- fix original 

## Notes

This note is created with Learn more about using [Distill for R Markdown](https://rstudio.github.io/distill).

As of now, source code is at [github.com/stared/bra-ket-vue-art](https://github.com/stared/bra-ket-vue-art).

This particualr post inspired by [Kronecker Product In Circuits by Lei Mao ](https://leimao.github.io/blog/Kronecker-Product-In-Circuits/) and [Quantum Entanglement and Superdense Coding by Victory Omole](https://vtomole.com/blog/2018/03/03/sd).


```{js, results='asis', echo=FALSE, message=FALSE}
const { Elements, Operator, Dimension, Cx } = QuantumTensors;
const { ComplexLegend, MatrixViewer } = BraKetVue;

const opNot = Operator.fromSparseCoordNames([
  ['1', '0', Cx(1)],
  ['0', '1', Cx(1)],
], [Dimension.position(2, 'bit')]);

new Vue({
  el: "#op-not",
  template: "<matrix-viewer :operator-raw='opNot' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
    	opNot
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-and",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['0', '01', Cx(1)],
          ['0', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-or",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-xor",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['0', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-nand",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['1', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['0', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-copy",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '0', Cx(1)],
          ['11', '1', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-first",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['0', '01', Cx(1)],
          ['1', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-second",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '00', Cx(1)],
          ['1', '01', Cx(1)],
          ['0', '10', Cx(1)],
          ['1', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-swap",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(1)],
          ['10', '01', Cx(1)],
          ['01', '10', Cx(1)],
          ['11', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-erase",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['0', '0', Cx(1)],
          ['0', '1', Cx(1)],
        ],
        [Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-xor-2",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['10', '01', Cx(0.5)],
          ['10', '10', Cx(0.5)],
          ['00', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['11', '01', Cx(0.5)],
          ['11', '10', Cx(0.5)],
          ['01', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
new Vue({
  el: "#op-cnot",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(1)],
          ['11', '01', Cx(1)],
          ['10', '10', Cx(1)],
          ['01', '11', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```


```{js, results='asis', echo=FALSE, message=FALSE}

new Vue({
  el: "#op-and-2",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
  	  // use outer product instead?
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['00', '01', Cx(0.5)],
          ['00', '10', Cx(0.5)],
          ['10', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['01', '01', Cx(0.5)],
          ['01', '10', Cx(0.5)],
          ['11', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}
const oneBitMix = Operator.fromSparseCoordNames([
    ['0', '0', Cx(0.5)],
    ['0', '1', Cx(0.5)],
    ['1', '0', Cx(0.5)],
    ['1', '1', Cx(0.5)],
  ],
  [Dimension.position(2, 'bit')],
  [Dimension.position(2, 'bit')])

new Vue({
  el: "#op-and-3",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['00', '00', Cx(0.5)],
          ['00', '01', Cx(0.5)],
          ['00', '10', Cx(0.5)],
          ['10', '11', Cx(0.5)],
          ['01', '00', Cx(0.5)],
          ['01', '01', Cx(0.5)],
          ['01', '10', Cx(0.5)],
          ['11', '11', Cx(0.5)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit')]
      ).outer(oneBitMix)
		}
  }
})
```

```{js, results='asis', echo=FALSE, message=FALSE}

new Vue({
  el: "#op-and-3-sol",
  template: "<matrix-viewer :operator-raw='operator' :dark-mode='false' :show-legend='false' />",
  components: {
    MatrixViewer
  },
  data() {
  	return {
      operator: Operator.fromSparseCoordNames([
          ['000', '000', Cx(1)],
          ['010', '010', Cx(1)],
          ['001', '100', Cx(1)],
          ['111', '110', Cx(1)],
          ['100', '001', Cx(1)],
          ['110', '011', Cx(1)],
          ['101', '101', Cx(1)],
          ['011', '111', Cx(1)],
        ],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit'), Dimension.position(2, 'bit')],
        [Dimension.position(2, 'bit'), Dimension.position(2, 'bit'), Dimension.position(2, 'bit')])
		}
  }
})
```